The effects of varying depth in cosmic shear surveys

### 2019-10-23 paperdraft_ver2.tex
added ver2, showing the alternative Fig 4

While the first method suffers from the fact that the power-law index only holds for sources of redshifts $1\lesssim z \lesssim 2$, the second method is sensitive to the angular range $\theta_{\rm{ref}}$, at which the shear correlation functions are evaluated, which is fairly arbitrary. In our case, we chose $\theta_{\rm{ref}}\approx 11'$, which is roughly in the logarithmic middle between the range of the correlation functions, $[0.\!'5,300']$. The choice of other values for $\theta_{\rm{ref}}$ leads to a different amplitude of the change $\xi_\pm/\xi_\pm^{\rm{obs}}$, but does not affect its shape. Generally, a smaller $\theta_{\rm{ref}}$ leads to a stronger effect, in particular, the highest amplitude of the change is at $\theta_{\rm{ref}}=0.\!'5$.

@@

The results can be seen in Fig.~\ref{fig:all_xis}.
  \begin{figure*}
  \centering
  \includegraphics[width=0.9\linewidth]{images/xis_all1.pdf}
  \caption{The ratio of correlation functions measured for a uniform depth survey, $\xi_\pm$, and a KiDS-like variable depth survey, $\xi_\pm^{\rm obs}$, cross-correlating five tomographic bins (as denoted in the upper left corner of each panel). Results from mock KiDS-like data (red) can be compared to analytic models from Sect.~2 (average redshift weighting in green-solid, and auto-correlation $\xi_+$ weighting in green-dashed) and the semi-analytic model from Sect.~4 (blue). Mock data is limited to the angular regime which is not significantly impacted by resolution effects.The upper left triangle depicts the ratios of $\xi_+$, whereas the lower right triangle depicts the ratios of $\xi_-$. The semi-analytic model was computed for the best-fit cosmology of H18. For the characteristic scale $\theta_{\rm{ref}}$ of the second analytic method we chose $\theta_{\rm{ref}} \approx 11'$. Note that the seemingly better agreement between the mocks and the analytic method is purely coincidental; as the mocks only take galaxies with $z<2$ into account, they slightly underestimate the effect.}
  \label{fig:all_xis}
  \end{figure*}
We observe that for high-redshift bins, all methods yield consistent results. For low-redshift bins, there are discrepancies between the different models. However, the average-redshift weighted analytic model is only valid for high redshifts, whereas the auto-correlation $\xi_+$ weighted model is entirely dependent on the choice of $\theta_{\rm{ref}}$. For $\theta_{\rm{ref}}=11'$, as chosen here, both analytic models agree very well for all redshift bins. For $\theta_{\rm{ref}}=0.\!'5$ (not shown), the auto-correlation weighted method agrees with the semi-analytic one pretty well for $\xi_+$. As $\xi_-$ is affected much stronger by this effect\footnote{The effect on $\xi_-$ is much stronger due to the fact that in Equation \eqref{eq:xipm-pkappa}, $\xi_+$ is computed by filtering the power spectrum with the 0-th order Bessel function. This function peaks at $\ell\theta=0$, meaning that for all values of $\theta$, the correlation function $\xi_+$ is sensitive to small values of $\ell$, corresponding to large separations $\theta$. However, $\xi_-$ is obtained by filtering with the 4-th order Bessel function, which peaks at approximately $\ell\theta\approx 5$, so for different $\theta$ this function is sensitive to varying parts of the convergence power spectrum. A more detailed analysis of this can be found in the Appendix of \citet{2017MNRAS.471.4412K}.}, the analytic method is not able to trace this change for any choice of $\theta_{\rm{ref}}$. Furthermore we note that the effect seems to be strongest in the first redshift bin, which is not surprising, as there the average redshift between pointings varies the most (compare Fig. \ref{fig:nz_of_meanz}).

@@

In general, the semi-analytic model predicts a stronger effect than the mocks. This is due to the fact that the mocks are subject to a redshift-cutoff at $z=2$, meaning that they do not take the high-redshift tail of galaxies into account. This is of particular importance in the first redshift bin, where this feature is especially pronounced (compare Fig.~\ref{fig:nofz_quantiles}). When the same cutoff is applied to the models, the agreement on sub-pointing scales is striking (compare Fig.~\ref{fig:redshift_cutoff} in the Appendix). In particular this means that the mocks slightly underestimate the effect of varying depth.

@@

% \clearpage
\section{Additional Figures and Tables}
\begin{table}[!htb]
\centering
\caption{Limiting magnitudes for the ten quantiles}
\label{tab:maglim}
\begin{tabular}{r c}
\hline\hline
quantile & $r$-band depth \\
\hline
0 & 25.76 \\
10 & 26.06 \\
20 & 26.15 \\
30 & 26.19 \\
40 & 26.23 \\
50 & 26.27 \\
60 & 26.31 \\
70 & 26.34 \\
80 & 26.39 \\
90 & 26.44 \\
100 & 26.60 \\
\hline
\end{tabular}
\end{table}
\begin{figure}[!htb]
\centering
\includegraphics[width=0.9\linewidth]{images/xis_all_cutoff.pdf}
\caption{The same Figure as Fig.~\ref{fig:all_xis}, but in this case the redshift distributions $p_n^i(z)$ were set equal to zero for $z>2$ in the models to mimick the redshift-cutoff in the mocks. Any residual differences are explained by the sub-pointing depth variations that are accounted for in the mocks, but not in the models. These variations slightly reduce the effect of varying depth.}
\label{fig:redshift_cutoff}
\end{figure}
\begin{figure}[!htb]
\centering
\includegraphics[width=0.9\linewidth]{images/compare_cosmos.pdf}
\caption{Correction to the correlation functions in varying cosmologies. Depicted here are three flat sample cosmologies, where values within the 98\% CL of the KV450 survey were sampled.}
\label{fig:comparecosmo}
\end{figure}
\end{appendix}
\end{document}
%

>>>>

While the first method suffers from the fact that the power-law index only holds for sources of redshifts $1\lesssim z \lesssim 2$, the second method is sensitive to the angular range $\theta_{\rm{ref}}$, at which the shear correlation functions are evaluated, which is fairly arbitrary. For $\theta_{\rm{ref}}\approx 11'$, which is roughly in the logarithmic middle between the range of the correlation functions, $[0.\!'5,300']$, the two calibration methods agree pretty well. The choice of other values for $\theta_{\rm{ref}}$ leads to a different amplitude of the change $\xi_\pm/\xi_\pm^{\rm{obs}}$, but does not affect its shape. Generally, a smaller $\theta_{\rm{ref}}$ leads to a stronger effect, in particular, the highest amplitude of the change is at $\theta_{\rm{ref}}=0.\!'5$.

@@

The results can be seen in Fig.~\ref{fig:all_xis}.
  \begin{figure*}
  \centering
  \includegraphics[width=0.9\linewidth]{images/xis_all2.pdf}
  \caption{The ratio of correlation functions measured for a uniform depth survey, $\xi_\pm$, and a KiDS-like variable depth survey, $\xi_\pm^{\rm obs}$, cross-correlating five tomographic bins (as denoted in the upper left corner of each panel). The upper left triangle depicts the ratios of $\xi_+$, whereas the lower right triangle depicts the ratios of $\xi_-$. Results from mock KiDS-like data (red) can be compared to analytic models from Sect.~2 (average redshift weighting, green), and the semi-analytic model from Sect.~3 (blue solid). Mock data is limited to the angular regime which is not significantly impacted by resolution effects. As the mocks only take galaxies with $z<2$ into account, they slightly underestimate the effect. Applying the same redshift-cutoff to the semi-analytical model (blue dashed) yields a near-perfect agreement on sub-pointing scales. Therefore, the seemingly better agreement between the mocks and the analytic method is purely coincidental. The semi-analytic model was computed for the best-fit cosmology of H18.}
  \label{fig:all_xis}
  \end{figure*}
We observe that for high-redshift bins, all methods yield consistent results. For low-redshift bins, there are discrepancies between the different models. However, the average-redshift weighted analytic model, as shown here, is only valid for high redshifts, whereas the auto-correlation $\xi_+$ weighted model (not shown) is entirely dependent on the choice of $\theta_{\rm{ref}}$. For $\theta_{\rm{ref}}=11'$, both analytic models agree very well for all redshift bins. For $\theta_{\rm{ref}}=0.\!'5$, the auto-correlation weighted method agrees with the semi-analytic one pretty well for $\xi_+$. As $\xi_-$ is affected much stronger by this effect\footnote{The effect on $\xi_-$ is much stronger due to the fact that in Equation \eqref{eq:xipm-pkappa}, $\xi_+$ is computed by filtering the power spectrum with the 0-th order Bessel function. This function peaks at $\ell\theta=0$, meaning that for all values of $\theta$, the correlation function $\xi_+$ is sensitive to small values of $\ell$, corresponding to large separations $\theta$. However, $\xi_-$ is obtained by filtering with the 4-th order Bessel function, which peaks at approximately $\ell\theta\approx 5$, so for different $\theta$ this function is sensitive to varying parts of the convergence power spectrum. A more detailed analysis of this can be found in the Appendix of \citet{2017MNRAS.471.4412K}.}, the analytic method is not able to trace this change for any choice of $\theta_{\rm{ref}}$. Furthermore we note that the effect seems to be strongest in the first redshift bin, which is not surprising, as there the average redshift between pointings varies the most (compare Fig. \ref{fig:nz_of_meanz}).

@@

In general, the semi-analytic model predicts a stronger effect than the mocks. This is due to the fact that the mocks are subject to a redshift-cutoff at $z=2$, meaning that they do not take the high-redshift tail of galaxies into account. This is of particular importance in the first redshift bin, where this feature is especially pronounced (compare Fig.~\ref{fig:nofz_quantiles}). When the same cutoff is applied to the models, the agreement on sub-pointing scales is striking. In particular this means that the mocks slightly underestimate the effect of varying depth.

@@

% \clearpage
\section{Additional Figures and Tables}
\begin{table}[!htb]
\centering
\caption{Limiting magnitudes for the ten quantiles}
\label{tab:maglim}
\begin{tabular}{r c}
\hline\hline
quantile & $r$-band depth \\
\hline
0 & 25.76 \\
10 & 26.06 \\
20 & 26.15 \\
30 & 26.19 \\
40 & 26.23 \\
50 & 26.27 \\
60 & 26.31 \\
70 & 26.34 \\
80 & 26.39 \\
90 & 26.44 \\
100 & 26.60 \\
\hline
\end{tabular}
\end{table}
\begin{figure}[!htb]
\centering
\includegraphics[width=0.9\linewidth]{images/compare_cosmos.pdf}
\caption{Correction to the correlation functions in varying cosmologies. Depicted here are three flat sample cosmologies, where values within the 98\% CL of the KV450 survey were sampled.}
\label{fig:comparecosmo}
\end{figure}
\end{appendix}
\end{document}
%
